Untitled

### 2025-03-14 script/r/discontinuity.Rmd
maj

```{r}
# Data manipulation and utility
library(dplyr)      # Data manipulation
library(readr)      # Data import/export
library(magrittr)   # Pipe operator (%>%)

# Statistical analysis and modeling
library(estimatr)   # Robust regression
library(car)        # Companion to Applied Regression (e.g., recoding variables)

# Visualization
library(ggplot2)    # Plotting
library(texreg)     # Exporting regression results to LaTeX/HTML

# Spatial and geospatial analysis
library(sf)         # Simple Features for R
library(geosphere)  # Geospatial functions
library(rnaturalearth)     # Natural Earth map data
library(rnaturalearthdata) # Data for rnaturalearth

# Specialized packages
library(rdrobust)   # Regression discontinuity designs
library(murdock)    # Specific package, assuming you have a need for it

library(haven)

setwd("~/working_paper/trust/")
getwd()
```

```{r}
# Load the Murdock ethnic groups map (1956)
murdock <- murdock::murdock

# Convert the Murdock map to an 'sf' (Simple Features) object for spatial data manipulation
murdock_sf <- st_as_sf(murdock)

# Define a set of African countries of interest by specifying their names
countries <- ne_countries(country = c("Benin","Botswana","Burkina Faso","Cameroon","Ivory Coast",
                                    "Guinea","Kenya","Liberia","Malawi","Mali",
                                    "Mozambique","Namibia","Niger","Nigeria","Sierra Leone",
                                    "United Republic of Tanzania","Uganda","Zambia","Zimbabwe",
                                    "Ghana"))

# Convert the selected countries into an 'sf' object for spatial operations
countries_sf <- st_as_sf(countries)

# Extract the individual boundaries of the selected African countries
boundaries <- st_boundary(countries_sf)

@@

ggsave("plots/discontinuity/murdock_map.jpg", width = 10, height = 6, dpi = 300)

# Create a 50km buffer around the African country borders
buffer <- st_buffer(boundaries, dist = 25000)

# Plot the Murdock map with the 25km buffers around the African borders
ggplot() +
  geom_sf(data = murdock_sf, fill = "transparent", color = "black", size = 1) + # Murdock map in black
  geom_sf(data = buffer, color = "red", lwd = 1.5) + # 25km buffer around borders in red
  theme_minimal() +
  ggtitle("25km Buffer of African Borders and Murdock Ethnic Zones (1956)")
```

```{r}
# List of countries of interest
countries <- c("Benin","Botswana","Burkina Faso","Cameroon","Ivory Coast",
                                    "Guinea","Kenya","Liberia","Malawi","Mali",
                                    "Mozambique","Namibia","Niger","Nigeria","Sierra Leone",
                                    "United Republic of Tanzania","Uganda","Zambia","Zimbabwe",
                                    "Ghana")

# Buffer distances
buffer_distances <- c(50000, 40000, 30000, 25000, 20000, 15000, 10000)

@@

# List of objects to keep
objects_to_keep <- c("data_trust_test", "union_list", "murdock_sf", "countries", "countries_sf", "buffer", "buffer_distances")

# Get all objects in the current environment

# Determine which objects to remove

# Remove the unnecessary objects

@@

# Example of plotting the 25km buffer result
ggplot() +
  geom_sf(data = murdock_sf, fill = "transparent", color = "black", size = 1) +
  geom_sf(data = unions_list[["20000"]], color = "red", lwd=1, fill="transparent") +  # 25km buffer
  theme_minimal(base_size = 15) +
  theme(
    panel.grid.major = element_blank(),  
    panel.grid.minor = element_blank(),  
    panel.background = element_blank(),   
    axis.text = element_blank(),         
    axis.ticks = element_blank(),         
    axis.title = element_blank(),         
    legend.key = element_rect(fill = "transparent", color = NA),
    legend.text = element_text(size = 8),
    legend.position = c(0.25, 0.35)
  )

ggsave("buffer_20.jpg", width = 10, height = 6, dpi = 300)

# List of objects to keep
objects_to_keep <- c("data_trust_test", "unions_list", "murdock_sf")

# Get all objects in the current environment

# Determine which objects to remove

# Remove the unnecessary objects

@@

```{r}
# Load data
data_trust_test <- readRDS("data/data_rds/data_final_r.rds")
bdd_geo <- read_csv("data/data_raw/bdd_geo_w8.csv")

# Merge and clean data
data_trust_test %<>%
  left_join(bdd_geo, by = "id_unique") %>%
  select(-...1) %>%
  mutate(
    distance_border_km = distance_border / 1000,
    distance_border_km2 = distance_border_km^2,
    distance_border_km3 = distance_border_km^3
  )

# Convert to spatial data and filter by borders
border_sample_50_union <- unions_list[["50000"]]
border_sample_40_union <- unions_list[["40000"]]
border_sample_30_union <- unions_list[["30000"]]
border_sample_25_union <- unions_list[["25000"]]
border_sample_20_union <- unions_list[["20000"]]
border_sample_15_union <- unions_list[["15000"]]
border_sample_10_union <- unions_list[["10000"]]

@@

# Filter data
filtered_data_list <- filter_by_border(data_trust_test_sf, borders)
names(filtered_data_list) <- c("50", "40", "30", "25", "20", "15", "10")

# Remove border polygons
rm(list = names(borders))

@@

# List of objects to keep
objects_to_keep <- c("df_list", "data_trust_test")

# Get all objects in the current environment

# Determine which objects to remove

# Remove the unnecessary objects

@@

capitals <- data_trust_test %>%
  select(caplong, caplat, iso) %>%
  na.omit() %>%
  rename(longitude = caplong, latitude = caplat)

@@

ggsave("plots/map/obs_map.jpg", width = 10, height = 10, units = "in", dpi = 300)

#selected_countries <- countries %>% 
 # filter(iso_a3 %in% c("BEN", "NER"))

#selected_individuals <- individuals %>%
 # filter(iso%in% c("BEN", "NER"))

#selected_capitals <- capitals %>%
 # filter(iso%in% c("BEN", "NER"))

#test <- murdock_sf%>%
#  distinct(NAME)

#selected_murdock <- murdock_sf%>%
#  filter(NAME=="DENDI")

#test <- unions_list[["20000"]]

#test %<>%
 # filter(iso%in%c("BEN", "NER"))

# Create the plot
#ggplot() +
  # Base map with country borders
#    geom_sf(data = selected_countries, fill = "white", color = "black") +
#  geom_sf(data = selected_murdock, fill = "green", color = "darkgreen", size = 1, alpha=0.5) +
  # Use theme_void to create a clean background
 # theme_void() + 
  # Adding capital cities with refined aesthetics
#  geom_point(data = selected_capitals, 
   #          aes(x = longitude, y = latitude, color = "Capitals"), 
  #           size = 2, shape = 22, fill = "red", show.legend = FALSE) +

#geom_label(aes(x = 2.5, y = 10, label = "Benin"), color = "black", size = 6, fill = "lightgrey", label.size = 0.5) +
 #     geom_label(aes(x = 4, y = 14.5, label = "Niger"), color = "black", size = 6, fill = "lightgrey", label.size = 0.5) +

#geom_text(aes(x = 2.4, y = 13.75, label = "Niamey"), color = "black", size = 5) +
#geom_text(aes(x = 2.2, y = 6.6, label = "Porto-Novo"), color = "black", size = 5) +

#  labs(color = NULL) +
 #   coord_sf(xlim = c(0, 5), ylim = c(6, 18), expand = FALSE)

# Save the plot if needed
#ggsave("writing/presentation/bdd_map.jpg", width = 10, height = 10, units = "in", dpi = 300)

# List of objects to keep
objects_to_keep <- c("df_list")

# Get all objects in the current environment

# Determine which objects to remove

# Remove the unnecessary objects

@@

# Apply the function to each data frame in the list with respective thresholds
df_list <- lapply(names(df_list), function(name) {
  df <- df_list[[name]]
  threshold <- case_when(
  name == "50" ~ 50000,
  name == "40" ~ 40000,
  name == "30" ~ 30000,
  name == "25" ~ 25000,
  name == "20" ~ 20000,
  name == "15" ~ 15000,
  name == "10" ~ 10000,
  TRUE ~ NA_real_
)
  df_filtered <- process_df(df, threshold)
  return(df_filtered)
})

# Update the names of the list to match the original names
names(df_list) <- c("df_50", "df_40", "df_30", "df_25", "df_20", "df_15", "df_10")
```

@@

# Apply the function to each data frame in the list
df_list <- lapply(df_list, convert_labelled_to_factors)

# Optionally, update list names if needed
names(df_list) <- c("df_50", "df_40", "df_30", "df_25", "df_20", "df_15", "df_10")

rm("convert_labelled_to_factors", "process_df")
```

@@

# Handle list columns by converting them to strings (or you could flatten or remove them as needed)
df_50[list_columns] <- lapply(df_50[list_columns], function(x) sapply(x, toString))

@@

write_dta(df_50, "data/data_dta/df_50.dta")

df_50 %<>%
  group_by(murdock_round) %>%
  filter(n_distinct(iso) >= 2)%>%
ungroup()

saveRDS(df_50, file="data/data_rds/df_50_2.rds")

write_dta(df_50, "data/data_dta/df_50_2.dta")

@@

# Handle list columns by converting them to strings (or you could flatten or remove them as needed)
df_40[list_columns] <- lapply(df_40[list_columns], function(x) sapply(x, toString))

@@

write_dta(df_40, "data/data_dta/df_40.dta")

df_40 %<>%
  group_by(murdock_round) %>%
  filter(n_distinct(iso) >= 2)%>%
ungroup()

saveRDS(df_40, file="data/data_rds/df_40_2.rds")

write_dta(df_40, "data/data_dta/df_40_2.dta")
```

@@

# Handle list columns by converting them to strings (or you could flatten or remove them as needed)
df_30[list_columns] <- lapply(df_30[list_columns], function(x) sapply(x, toString))

@@

saveRDS(df_30, file="data/data_rds/df_30.rds")
write_dta(df_30, "data/data_dta/df_30.dta")

df_30 %<>%
  group_by(murdock_round) %>%
  filter(n_distinct(iso) >= 2)%>%
ungroup()

saveRDS(df_30, file="data/data_rds/df_30_2.rds")

write_dta(df_30, "data/data_dta/df_30_2.dta")

@@

# Handle list columns by converting them to strings (or you could flatten or remove them as needed)
df_25[list_columns] <- lapply(df_25[list_columns], function(x) sapply(x, toString))

@@

write_dta(df_25, "data/data_dta/df_25.dta")

df_25 %<>%
  group_by(murdock_round) %>%
  filter(n_distinct(iso) >= 2)%>%
ungroup()

saveRDS(df_25, file="data/data_rds/df_25_2.rds")

write_dta(df_25, "data/data_dta/df_25_2.dta")
```

@@

# Handle list columns by converting them to strings (or you could flatten or remove them as needed)
df_20[list_columns] <- lapply(df_20[list_columns], function(x) sapply(x, toString))

@@

write_dta(df_20, "data/data_dta/df_20.dta")

df_20 %<>%
  group_by(murdock_round) %>%
  filter(n_distinct(iso) >= 2)%>%
ungroup()

saveRDS(df_20, file="data/data_rds/df_20_2.rds")

write_dta(df_20, "data/data_dta/df_20_2.dta")
```

@@

# Handle list columns by converting them to strings (or you could flatten or remove them as needed)
df_15[list_columns] <- lapply(df_15[list_columns], function(x) sapply(x, toString))

@@

write_dta(df_15, "data/data_dta/df_15.dta")

df_15 %<>%
  group_by(murdock_round) %>%
  filter(n_distinct(iso) >= 2)%>%
ungroup()

saveRDS(df_15, file="data/data_rds/df_15_2.rds")

write_dta(df_15, "data/data_dta/df_15_2.dta")
```

@@

# Handle list columns by converting them to strings (or you could flatten or remove them as needed)
df_10[list_columns] <- lapply(df_10[list_columns], function(x) sapply(x, toString))

@@

write_dta(df_10, "data/data_dta/df_10.dta")

df_10 %<>%
  group_by(murdock_round) %>%
  filter(n_distinct(iso) >= 2)%>%
ungroup()

saveRDS(df_10, file="data/data_rds/df_10_2.rds")

write_dta(df_10, "data/data_dta/df_10_2.dta")
```

>>>>

```{r}
setwd("~/working_paper/trust/")
getwd()

library(dplyr)
library(readr)
library(magrittr)
library(ggplot2)
library(sf)
library(rnaturalearth)
library(murdock)
library(haven)
```

```{r}
murdock <- murdock::murdock
murdock_sf <- st_as_sf(murdock)

countries <- ne_countries(country = c("Benin","Botswana","Burkina Faso","Cameroon","Ivory Coast",
                                    "Guinea","Kenya","Liberia","Malawi","Mali",
                                    "Mozambique","Namibia","Niger","Nigeria","Sierra Leone",
                                    "United Republic of Tanzania","Uganda","Zambia","Zimbabwe",
                                    "Ghana"))

countries_sf <- st_as_sf(countries)

boundaries <- st_boundary(countries_sf)

@@

ggsave("C:/Users/Redha CHABA/Documents/wp_git/rbci/plots/discontinuity/murdock_map.jpg", width = 10, height = 6, dpi = 300)

buffer <- st_buffer(boundaries, dist = 25000)

ggplot() +
  geom_sf(data = murdock_sf, aes (color="Murdock"), fill = "transparent", size = 1) + # Murdock map in black
  geom_sf(data = buffer, aes(color = "boundaries"), lwd = 1) + # 25km buffer around borders in red
  scale_color_manual(
    values = c("Murdock" = "black", "boundaries" = "red"),
    labels = c("25km boundaries", "Historical Ethnic Areas"),
    name = NULL 
  ) +
  theme_minimal(base_size = 15) +
  theme(
    panel.grid.major = element_blank(),  
    panel.grid.minor = element_blank(),  
    panel.background = element_blank(),   
    axis.text = element_blank(),         
    axis.ticks = element_blank(),         
    axis.title = element_blank(),         
    legend.key = element_rect(fill = "transparent", color = NA),
    legend.text = element_text(size = 8),
    legend.position = c(0.25, 0.35)
  )

ggsave("C:/Users/Redha CHABA/Documents/wp_git/rbci/plots/discontinuity/murdock_boundaries_map.jpg", width = 10, height = 6, dpi = 300)
```

```{r}
countries <- c("Benin","Botswana","Burkina Faso","Cameroon","Ivory Coast",
                                    "Guinea","Kenya","Liberia","Malawi","Mali",
                                    "Mozambique","Namibia","Niger","Nigeria","Sierra Leone",
                                    "United Republic of Tanzania","Uganda","Zambia","Zimbabwe",
                                    "Ghana")

buffer_distances <- c(50000, 40000, 30000, 25000, 20000, 15000, 10000)

@@

@@

# Example of plotting the 25km buffer result
ggplot() +
  geom_sf(data = murdock_sf, fill = "transparent", aes(color = "Murdock"), size = 1) +
  geom_sf(data = unions_list[["25000"]], aes(color = "boundaries"), lwd=1, fill="transparent") +
  scale_color_manual(
    values = c("Murdock" = "black", "boundaries" = "red"),
    labels = c("25km boundaries", "Historical Ethnic Areas"),
    name = NULL 
  ) +
  theme_minimal(base_size = 15) +
  theme(
    panel.grid.major = element_blank(),  
    panel.grid.minor = element_blank(),  
    panel.background = element_blank(),   
    axis.text = element_blank(),         
    axis.ticks = element_blank(),         
    axis.title = element_blank(),         
    legend.key = element_rect(fill = "transparent", color = NA),
    legend.text = element_text(size = 8),
    legend.position = c(0.25, 0.35)
  )

ggsave("buffer_25.jpg", width = 10, height = 6, dpi = 300)

@@

```{r}
data_trust_test <- readRDS("data/data_rds/data_27_02_25.rds")
data_trust_test%<>%
  ungroup()

bdd_geo <- read_csv("data/data_raw/bdd_geo_w8.csv")

data_trust_test %<>%
  filter(spl==1)%>%
  left_join(bdd_geo, by = "id_unique") %>%
  select(-...1) %>%
  mutate(
    distance_border_km = distance_border / 1000,
    distance_border_km2 = distance_border_km^2,
    distance_border_km3 = distance_border_km^3
  )

border_sample_50_union <- unions_list[["50000"]]
border_sample_40_union <- unions_list[["40000"]]
border_sample_30_union <- unions_list[["30000"]]
border_sample_25_union <- unions_list[["25000"]]
border_sample_20_union <- unions_list[["20000"]]
border_sample_15_union <- unions_list[["15000"]]
border_sample_10_union <- unions_list[["10000"]]

@@

# Filter data
filtered_data_list <- filter_by_border(data_trust_test_sf, borders)

names(filtered_data_list) <- c("50", "40", "30", "25", "20", "15", "10")
sapply(filtered_data_list, nrow)

@@

@@

capitals <- data_trust_test %>%
  distinct(caplong, caplat, iso)%>%
  na.omit() %>%
  rename(longitude = caplong, latitude = caplat)

@@

ggsave("plots/map/obs_map.jpg", width = 10, height = 10, units = "in", dpi = 300)
```

@@

df_list <- lapply(names(df_list), function(name) {
  df <- df_list[[name]]
  threshold <- case_when(
  name == "50" ~ 50000,
  name == "40" ~ 40000,
  name == "30" ~ 30000,
  name == "25" ~ 25000,
  name == "20" ~ 20000,
  name == "15" ~ 15000,
  name == "10" ~ 10000,
  TRUE ~ NA_real_
)
  df_filtered <- process_df(df, threshold)
  return(df_filtered)
})

names(df_list) <- c("df_50", "df_40", "df_30", "df_25", "df_20", "df_15", "df_10")
```

@@

df_list <- lapply(df_list, convert_labelled_to_factors)

names(df_list) <- c("df_50", "df_40", "df_30", "df_25", "df_20", "df_15", "df_10")
```

@@

df_50[list_columns] <- lapply(df_50[list_columns], function(x) sapply(x, toString))

@@

write_dta(df_50, "data/data_dta/df_50.dta")
```

@@

df_40[list_columns] <- lapply(df_40[list_columns], function(x) sapply(x, toString))

@@

write_dta(df_40, "data/data_dta/df_40.dta")
```

@@

df_30[list_columns] <- lapply(df_30[list_columns], function(x) sapply(x, toString))

@@

saveRDS(df_30, file="data/data_rds/df_30.rds")
write_dta(df_30, "data/data_dta/df_30.dta")
```

@@

df_25[list_columns] <- lapply(df_25[list_columns], function(x) sapply(x, toString))

@@

write_dta(df_25, "data/data_dta/df_25.dta")
```

@@

df_20[list_columns] <- lapply(df_20[list_columns], function(x) sapply(x, toString))

@@

write_dta(df_20, "data/data_dta/df_20.dta")
```

@@

df_15[list_columns] <- lapply(df_15[list_columns], function(x) sapply(x, toString))

@@

write_dta(df_15, "data/data_dta/df_15.dta")
```

@@

df_10[list_columns] <- lapply(df_10[list_columns], function(x) sapply(x, toString))

@@

write_dta(df_10, "data/data_dta/df_10.dta")
```

### 2025-03-14 script/r/discontinuity_plot.Rmd
maj

  # Left side smooth curve (Close to Capital)
# Left side smooth curve (Close to Capital)
geom_smooth(data = data_left_plus,
            aes(x = mean_dist_border_intervals, y = mean_trust_intervals),
            method = "lm", formula = y ~ poly(x, 3), 
            color = "navy", size = 1.2, se = TRUE,
            fill="navy",
            alpha=0.1,
            level = 0.90, show.legend = FALSE) +

# Right side smooth curve (Far from Capital)
geom_smooth(data = data_right_plus,
            aes(x = mean_dist_border_intervals, y = mean_trust_intervals),
            method = "lm", formula = y ~ poly(x, 3), 
            color = "firebrick", size = 1.2, se = TRUE,
            fill = "firebrick",
            alpha = 0.1,
            level = 0.90, show.legend = FALSE)+

geom_segment(aes(x = -40000, 
                   xend = 0,
                   y = moy_trust_pres, 
                   yend = moy_trust_pres, linetype = "Close"),
             inherit.aes = FALSE,
             color = "navy", size = 1.5, alpha = 0.7, data = data.frame()) + 
  geom_segment(aes(x = 0, 
                   xend =40000, 
                   y = moy_trust_loin, 
                   yend = moy_trust_loin, linetype = "Far"),
               inherit.aes = FALSE,
               color = "firebrick", size = 1.5, alpha = 0.7, data = data.frame()) +

@@

  annotate("label", x = -max(abs(data_intervals$mean_dist_border_intervals)) * 0.5, 
           y = max(data_intervals_ic$mean_trust_intervals) + 0.05, 
           label = "Close to the Capital", 
           size = 7, fill = "aliceblue", color = "navy", fontface = "bold", 
           label.size = 0.5, family = "Arial") +
  annotate("label", x = max(abs(data_intervals$mean_dist_border_intervals)) * 0.5, 
           y = max(data_intervals_ic$mean_trust_intervals) + 0.05, 
           label = "Far from the Capital", 
           size = 7, fill = "mistyrose", color = "firebrick", fontface = "bold", 
           label.size = 0.5, family = "Arial") +

  # Add legend for mean lines
  scale_linetype_manual(name = NULL, 
                        values = c("Far" = "dashed", 
                                  "Close" = "dashed"),
                        labels = c(expression(bar(y)[far]), expression(bar(y)[close]))) +

@@

ggsave("C:/Users/Redha CHABA/Documents/wp_git/rbci/plots/discontinuity/discontinuity.jpg", plot = last_plot(), width = 10, height = 6, units = "in", dpi = 300)
```

>>>>

  # Left side smooth curve (Close to Capital)
# Left side smooth curve (Close to Capital)
geom_smooth(data = data_left_plus,
            aes(x = mean_dist_border_intervals, y = mean_trust_intervals),
            method = "lm", formula = y ~ poly(x, 3), 
            color = "firebrick", size = 1.2, se = TRUE,
            fill="firebrick",
            alpha=0.1,
            level = 0.90, show.legend = FALSE) +

# Right side smooth curve (Far from Capital)
geom_smooth(data = data_right_plus,
            aes(x = mean_dist_border_intervals, y = mean_trust_intervals),
            method = "lm", formula = y ~ poly(x, 3), 
            color = "darkgreen", size = 1.2, se = TRUE,
            fill = "darkgreen",
            alpha = 0.1,
            level = 0.90, show.legend = FALSE)+

geom_segment(aes(x = -40000, 
                   xend = 0,
                   y = moy_trust_pres, 
                   yend = moy_trust_pres, linetype = "Close"),
             inherit.aes = FALSE,
             color = "firebrick", size = 1.5, alpha = 0.7, data = data.frame()) + 
  geom_segment(aes(x = 0, 
                   xend =40000, 
                   y = moy_trust_loin, 
                   yend = moy_trust_loin, linetype = "Far"),
               inherit.aes = FALSE,
               color = "darkgreen", size = 1.5, alpha = 0.7, data = data.frame()) +

@@

  annotate("label", x = -max(abs(data_intervals$mean_dist_border_intervals)) * 0.5, 
           y = max(data_intervals_ic$mean_trust_intervals) + 0.05, 
           label = "Close to the Capital", 
           size = 7, fill = "mistyrose", color = "firebrick", fontface = "bold", 
           label.size = 0.5, family = "Arial") +
  annotate("label", x = max(abs(data_intervals$mean_dist_border_intervals)) * 0.5, 
           y = max(data_intervals_ic$mean_trust_intervals) + 0.05, 
           label = "Far from the Capital", 
           size = 7, fill = "honeydew", color = "darkgreen", fontface = "bold", 
           label.size = 0.5, family = "Arial") +

  # Add legend for mean lines
  scale_linetype_manual(name = NULL, 
                        values = c("Close" = "dashed","Far" = "dashed"),
                        labels = c(expression(bar(y)[far]), expression(bar(y)[close]))) +

@@

ggsave("C:/Users/Redha CHABA/Documents/wp_git/rbci/plots/discontinuity/discontinuity2.jpg", plot = last_plot(), width = 10, height = 6, units = "in", dpi = 300)
```
